Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2015-11-05 WSSSPE3/report/WSSSPE3_review.tex
more small edits in lightning talk section

\begin{enumerate}
\item \textbf{Benjamin Tovar and Douglas Thain: \textit{Freedom vs. Stability:
Facilitating Research Training While Supporting Scientific Research}}
A case study: ``The Cooperative Computing Lab'' at the University of Notre Dame
is a small group of individuals with the main tasks of collaborating with
people that have large-scale computing problems, operating various parallel
computer systems, conducting computer science research, and developing open
source software. One of the main challenges they face is finding the balance
between flexibility/training and stability/quality. Their current solution for
ensuring the latter was to add a software engineer (the presenter) to the
existing team of faculty and students, who now also serves as ``spring''
between flexibility and stability.

\item \textbf{Birgit Penzenstadler, Colin Venters, Christoph Becker, Stefanie
Betz, Ruzanna Chitchyan, Letícia Duboc, Steve Easterbrook, Guillermo
Rodriguez-Navas and Norbert Seyff: \textit{Manifesting the Ghost of the Future:
Sustainability}}
The concept of sustainability has become a topic of interest in the field of
computing, which is evidenced by the increase in the number of events that
focus on the topic - nevertheless, it isn’t well understood yet. Penzenstadler
argued that we often define sustainability too narrowly. Instead,
sustainability at its heart is a systemic concept and must be viewed from a
range of different dimensions including environmental, economic, individual,
social and technical. She introduced the Karlskrona Manifesto on Software
Design, which distills knowledge from a broad range of related work on the
topic of sustainability into a set of (mis-)perceptions and principles. The
manifesto does not proclaim that there is an easy, one size fits all solution
around the corner, but rather points out that sustainability is a ``wicked
problem'' and is often misunderstood. Due to these misperceptions, even though
sustainability's importance is increasingly recognized, many software systems
are unsustainable. Even more alarming is that most software systems' broader
impacts on sustainability are unknown. To change this, the Karlskrona Manifesto
proposes nine principles and commitments. These commitments are not dogmatic
laws, but rather a commitment to rethink, to move beyond the silo mentality,
and analyze in more depth. As such, they do not restrict, but rather open up a
space for discussion.

>>>>

\begin{enumerate}
\item \textbf{Benjamin Tovar and Douglas Thain: \textit{Freedom vs. Stability:
Facilitating Research Training While Supporting Scientific Research}}
A case study: ``The Cooperative Computing Lab'' at the University of Notre Dame
is a small group of individuals whose main tasks are collaborating with
people that have large-scale computing problems, operating various parallel
computer systems, conducting computer science research, and developing open
source software. One of the main challenges they face is finding the balance
between flexibility/training and stability/quality. Their current solution for
ensuring the latter was to add a software engineer (the presenter) to the
existing team of faculty and students, who now also serves as ``spring''
between flexibility and stability.

\item \textbf{Birgit Penzenstadler, Colin Venters, Christoph Becker, Stefanie
Betz, Ruzanna Chitchyan, Let\'{i}cia Duboc, Steve Easterbrook, Guillermo
Rodriguez-Navas and Norbert Seyff: \textit{Manifesting the Ghost of the Future:
Sustainability}}
The concept of sustainability has become a topic of interest in the field of
computing, which is evidenced by the increase in the number of events that
focus on the topic - nevertheless, it isn't well understood yet. Penzenstadler
argued that we often define sustainability too narrowly. Instead,
sustainability at its heart is a systemic concept and must be viewed from a
range of different dimensions including environmental, economic, individual,
social and technical. She introduced the Karlskrona Manifesto on Software
Design~\cite{Becker:2014}, which distills knowledge from a broad range of related work on the
topic of sustainability into a set of (mis-)perceptions and principles. The
manifesto does not proclaim that there is an easy, one size fits all solution
around the corner, but rather points out that sustainability is a ``wicked
problem'' and is often misunderstood. Due to these misperceptions, even though
sustainability's importance is increasingly recognized, many software systems
are unsustainable. Even more alarming is that most software systems' broader
impacts on sustainability are unknown. To change this, the Karlskrona Manifesto
proposes nine principles and commitments. These commitments are not dogmatic
laws, but rather a commitment to rethink, to move beyond the silo mentality,
and analyze in more depth. As such, they do not restrict, but rather open up a
space for discussion.
